---
title: "csviSubtype"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# csviSubtype

`csviSubtype` implements a GWAS-anchored and vascular-centered framework for molecular subtype barcode discovery.

The current v0.1 package focuses on three core tasks:

1. building scRNA-derived subtype gene features from CSS, CIS, LRS, and DEG inputs,
2. fusing scRNA and GWAS gene-level evidence into subtype barcodes,
3. summarizing robustness under ablation settings.

## Installation
Install from GitHub:
```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("YuqianLii/csviSubtype")
```

For local development:
```{r eval=FALSE}
# install.packages(c("devtools", "data.table"))
devtools::load_all()
```

## Main functions

- `csvi_build_scrna_features()`
- `csvi_fuse_barcode()`
- `csvi_summarize_ablation()`

## Included example data

The package includes example input files under `extdata/`:

- `css_example.tsv`
- `cis_example.tsv`
- `lrs_example.tsv`
- `deg_example.tsv`
- `gwas_feature_example.tsv`

These files are intended to demonstrate the package workflow and output structure.

## Minimal example
```{r minimal-example, message=FALSE, warning=FALSE}
library(data.table)
library(csviSubtype)

css <- fread(system.file("extdata", "css_example.tsv", package = "csviSubtype"))
cis <- fread(system.file("extdata", "cis_example.tsv", package = "csviSubtype"))
lrs <- fread(system.file("extdata", "lrs_example.tsv", package = "csviSubtype"))
deg <- fread(system.file("extdata", "deg_example.tsv", package = "csviSubtype"))
gwas <- fread(system.file("extdata", "gwas_feature_example.tsv", package = "csviSubtype"))

scrna_res <- csvi_build_scrna_features(
  css = css,
  cis = cis,
  lrs = lrs,
  deg = deg,
  mode = "receptor"
)

fusion_res <- csvi_fuse_barcode(
  gwas = gwas,
  scrna_a1 = scrna_res$features_by_group$A1,
  scrna_b1 = scrna_res$features_by_group$B1
)

names(scrna_res)
names(fusion_res)
```

## Inspect scRNA feature results
```{r inspect-scrna}
head(scrna_res$features)
table(scrna_res$features$scrna_tier)
```

## Inspect fused subtype ledgers
```{r inspect-fusion}
nrow(fusion_res$A1$full)
nrow(fusion_res$A1$binary)
nrow(fusion_res$A1$strict)

nrow(fusion_res$B1$full)
nrow(fusion_res$B1$binary)
nrow(fusion_res$B1$strict)
```

## View top binary entries
```{r top-binary}
head(fusion_res$A1$binary[, .(
  gene_key, gwas_pp4_max, gwas_tier, q_sc, scrna_tier, zone_binary
)])

head(fusion_res$B1$binary[, .(
  gene_key, gwas_pp4_max, gwas_tier, q_sc, scrna_tier, zone_binary
)])
```
## Output structure

`csvi_build_scrna_features()` returns a list with:

- `css_group`
- `cis_all`
- `cis_top`
- `lrs_use`
- `features`
- `features_by_group`

`csvi_fuse_barcode()` returns a list with:

- `gwas_gene_summary`
- `A1`
- `B1`

Each subtype entry contains:

- `full`
- `binary`
- `strict`

## Notes
- `mode = "receptor"` is currently implemented in `v0.1`.
- `mode = "pathway"` is reserved for future expansion.
- Example files are provided for workflow demonstration and package testing.

## Citation
If you use `csviSubtype`, please cite the associated preprint/manuscript when available.
